Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-28 paper/paper.tex
Expand experiments section

We present a simple-yet-effective method for synthesizing loop-free programs
from specifications, using a novel combination of Bounded Model Checking and
explicit-state Model Checking.  The user provides a specification in plain
C, after which the synthesis process is fully automatic.

We demonstrate the effectiveness of our method by synthesizing several subtle
programs acting on machine integers, and a previously un-synthesised floating-point program.
\end{abstract}

@@

@@

\iffalse
Our rules for excluding non-canonical instructions are:

@@

\section{Experimental Results}
\label{sec:experiments}
We implemented our fully automatic synthesis procedure as the {\sc kalashnikov} tool.  The implementation consists
of around 600 lines of \newC and 600 lines of Python.  To evaluate the tool we used the 25 bitvector
programs from~\cite{brahma} and~\cite{brahma-icse}.  The majority of these are ``bit twiddling hacks'' taken from
Hacker's Delight~\cite{hackers-delight}.  The code we used to perform the experiments, along with the
benchmarks, is available at \texttt{http://www.cprover.org/kalashnikov}.  As a backend solver, we used
{\sc CBMC}~\cite{cbmc-website} at SVN revision r2777, with Glucose 2.2~\cite{glucose-paper} as the SAT solver.
We performed our experiments on a 4-core, 3.07\,GHz Xeon X5667 with 50\,GB of RAM.

To give a reference point for the efficiency of our tool, we present the results given for {\sc brahma}
on the same benchmarks, as reported in~\cite{brahma} and~\cite{brahma-icse}\footnote{Unfortunately
we were not able to obtain a copy of {\sc Brahma}, and so were unable to rerun the experiments ourselves.}.
These experiments were performed an 8-core,
1.86\,GHz Xeon with 4\,GB of RAM.  Although there are 49 results shown across the two tables, it should
be noted that there is a lot of overlap between the two sets of benchmarks and that there are only
29 unique benchmarks altogether.

We also ran {\sc AHA} on our benchmarks and found that on the small problems (those with 2-instruction solutions)
{\sc AHA} found a single correct solution in negligible time.  On the larger benchmarks, candidate solutions
were still found rapidly (around 10\,s for the 3-instruction programs) but the number of candidates was
too large to practically check them all.  For example, {\sc AHA} found 2284262 candidate
programs in 12.13\,s for p12.

@@

{\bf {\sc Brahma} and {\sc Kalashnikov} both synthesise minimal programs:}
This represents the majority of cases.  A statistical breakdown of this case is given in Fig~\ref{fig:stats-table}.
{\sc Kalashnikov} performs substantially better than {\sc Brahma} on these tests, and similarly to
semibiased {\sc Brahma} (an optimisation introduced in~\cite{brahma-icse}).  Program synthesis appears to
be CPU-bound, and so part of the performance difference may be explained by the faster CPU in the machine
used to run {\sc Kalashnikov}.

{\bf {\sc Kalashnikov} is unable to synthesise a program:}
This happens in 8 of the 29 cases.  In each case, {\sc Brahma} needs user guidance to synthesise the program.
The existence of instances that are beyond the reach of current automatic
methods, and thus require manual intervention, is to be expected.
% A total of 8 tests fall into this category.

{\bf {\sc Kalashnikov} synthesises a shorter program than {\sc Brahma}:}
This happens in 4 of the 29 cases.  For instance, the specification for p24 of Fig.~\ref{fig:icse-results-table} 
is a segment of obfuscated C code taken from a piece of malware.  The code uses an apparently unbounded loop
to multiply its input by the constant 45.  {\sc Brahma} synthesises an equivalent program consisting of
4 instructions, whereas {\sc Kalashnikov} synthesises the single instruction \texttt{mul x 45}.

To demonstrate the capabilities of \newC as an implementation language, we synthesise a
non-trivial floating-point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating-point numbers.  The original {\sc 2Sum} algorithm is given as Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}, shown in Alg.~\ref{alg:fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  {\sc 2Sum} was proved to be minimal in~\cite{fast2sum}, but
we believe our synthesis constitutes the first proof that {\sc Fast2Sum} is
minimal under the assumption that $a \ge b$.

@@

\begin{figure}[p]
\begin{center}
{\tiny
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma}.
 \label{fig:results-table}}
\end{figure}

\begin{figure}[p]
\begin{center}
{\tiny
\include{figures/icse_results_table}
}
\end{center}
\caption{Times for synthesis of machine integer benchmarks from~\cite{brahma-icse}.}
\label{fig:icse-results-table}
\end{figure}

>>>>

We present a simple-yet-effective method for synthesising loop-free programs
from specifications, using a novel combination of Bounded Model Checking and
explicit-state Model Checking.  The user provides a specification in plain
C, after which the synthesis process is fully automatic and is guaranteed to
synthesise a minimal program.

We demonstrate the effectiveness of our method by synthesising several subtle
programs acting on machine integers, de-obfuscating complex C code and
synthesising a previously un-synthesised floating-point program.
\end{abstract}

@@

\iffalse

@@

Our rules for excluding non-canonical instructions are:

@@

\section{Experimental Results}
\label{sec:experiments}
We implemented our fully automatic synthesis procedure as the {\sc kalashnikov} tool.  The implementation consists
of around 600 lines of \newC and 600 lines of Python.  To evaluate the tool we used the 25 bitvector
programs from~\cite{brahma} and~\cite{brahma-icse}.  The majority of these are ``bit twiddling hacks'' taken from
Hacker's Delight~\cite{hackers-delight}.  The code we used to perform the experiments, along with the
benchmarks, is available at \texttt{http://www.cprover.org/kalashnikov}.  As a backend solver, we used
{\sc CBMC}~\cite{cbmc-website} at SVN revision r3545, with Glucose 3.0~\cite{glucose-paper} as the SAT solver.
We performed our experiments on a 4-core, 2.40\,GHz Xeon E5-2665 with 32\,GB of RAM.

To give a reference point for the efficiency of our tool, we present the results given for {\sc brahma}
on the same benchmarks, as reported in~\cite{brahma} and~\cite{brahma-icse}\footnote{Unfortunately
we were not able to obtain a copy of {\sc Brahma}, and so were unable to rerun the experiments ourselves.}.
These experiments were performed on an 8-core,
1.86\,GHz Xeon with 4\,GB of RAM.  Although there are 49 results shown across the two tables, it should
be noted that there is a lot of overlap between the two sets of benchmarks and that there are only
29 unique benchmarks altogether.

We also ran the {\sc AHA} superoptimiser on our benchmarks and found that on the small problems
(those with 2-instruction solutions)
{\sc AHA} found a single correct solution in negligible time.  On the larger benchmarks, candidate solutions
were still found rapidly (around 10\,s for the 3-instruction programs) but the number of candidates was
too large to practically check them all.  For example, {\sc AHA} found 2284262 candidate
programs in 12.13\,s for p12.

@@

\paragraph{\bf {\sc Kalashnikov} synthesises a shorter program than {\sc Brahma}:}
This happens in 4 of the 29 cases.  This case is illustrated by benchmark p29, shown in Fig.~\ref{fig:obfuscated}.
The specification here is a piece of obfuscated code taken from the Conficker worm~\cite{conficker}, and
our goal is to synthesise an equivalent program which is easier to understand.  The obfuscated code
uses several tricks, including an apparently unbounded loop.  {\sc Brahma} is able to
produce an equivalent program consisting of four instructions using shifts and addition, whereas {\sc kalashnikov}
is able to produce the minimal program \verb|y * 45|.  It is also worth noting that the specification
fed to {\sc kalashnikov} was just the obfuscated code, with no further preprocessing needed.
\paragraph{\bf {\sc Kalashnikov} is unable to synthesise a program:}
This happens in 8 of the 29 cases.  In each case, {\sc Brahma} needs user guidance to synthesise the program.
The existence of instances that are beyond the reach of current automatic
methods, and thus require manual intervention, is to be expected.
% A total of 8 tests fall into this category.
\paragraph{\bf {\sc Brahma} and {\sc Kalashnikov} both synthesise minimal programs:}
In the remaining cases, both {\sc brahma} and {\sc kalashnikov} are able to produce
minimal programs.  We performed a Mann-Whitney U-test (see Fig.~\ref{fig:stats-table})
to determine whether there was a significant difference
in the speed of the two tools.  The test was unable to reject the null hypothesis
(that the tools are equally fast) at the p=0.05
level, i.e. there is no statistically significant difference in the speed of
the two tools on this set of benchmarks.

Obfuscated code

\begin{lstlisting}[frame=single,language=c]
int obfuscated (int y) {
 int a=1, b=0, z=1, c=0;
 while(1) {
  if (a == 0) { if (b == 0) { y=z+y; a =!a; b=!b;c=!c;
  if (!c) break;} else { z=z+y; a=!a; b=!b; c=!c;
  if (!c) break;} } else { if (b == 0) { z=y << 2;
  a=!a; } else { z=y << 3; a=!a; b=!b; } }
 }
 return z;
}
\end{lstlisting}
%\end{minipage}

\begin{minipage}[t]{.45\textwidth}
{\sc Brahma}'s output

\begin{lstlisting}[language=c,frame=single]
int recovered (int y) {
 int z = y << 2;
 y = z + y;
 z = y << 3
 y = z + y;
 return z;
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}[t]{.45\textwidth}
{\sc Kalashnikov}'s output
 \begin{lstlisting}[language=c,frame=single]
int recovered (int y) {
 return y*45;
}
\end{lstlisting}
\end{minipage}

 \caption{De-obfuscating  C code.
  \label{fig:obfuscated}}
\end{figure}

\subsection{Floating Point}
To demonstrate the capabilities of \newC as an implementation language, we synthesise a
non-trivial floating-point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating-point numbers.  The original {\sc 2Sum} algorithm is given as Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}, shown in Alg.~\ref{alg:fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  {\sc 2Sum} was proved to be minimal in~\cite{fast2sum}, but
we believe our synthesis constitutes the first proof that {\sc Fast2Sum} is
minimal under the assumption that $a \ge b$.

@@

\begin{figure}[p]
\begin{center}
{\small
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma} and~\cite{brahma-icse}.
 \label{fig:results-table}}
\end{figure}

%\begin{figure}[p]
%\begin{center}
%{\tiny
%\include{figures/icse_results_table}
%}
%\end{center}
%\caption{Times for synthesis of machine integer benchmarks from~\cite{brahma-icse}.}
%\label{fig:icse-results-table}
%\end{figure}
